---
title: "Unit 6: Classification Rules"
author: "Steven Ferguson"
date: "2022-09-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

mushrooms <- read.csv("mushrooms.csv", stringsAsFactors = TRUE)

# examine the structure of the data frame
str(mushrooms)

# drop the veil_type feature
mushrooms$veil_type <- NULL
# examine the class distribution
table(mushrooms$type)
```

"We will consider the 81214 samples in the mushroom data to be an exhaustive set of all the possible wild mushrooms

**IMPORTANT ASSUMPTION**

because this is a whole population, not a sample, we don't need to split data up into test and training datasets We are not trying to develop rules that cover unforeseen types of Mushrooms

ZeroR ignores all features and simply predict's target's MODE edible \> #poisonous, therefore it would state ALL mushrooms are edible.... not great

```{r}
#install.packages("OneR")
library(OneR)

# train OneR() on the data
mushroom_1R = OneR(type ~ ., data = mushrooms)

print(mushroom_1R)


```

```{r}
mushroom_1R_pred <- predict(mushroom_1R, mushrooms)
table(actual = mushrooms$type, predicted = mushroom_1R_pred)
# DANGEROUS RULE. TOO MANY FALSE POSITIVES. 120 predicted edible, but actually poisonous
```

```{r}
detach("package:OneR", unload = TRUE)
#install.packages("RWeka")
library(RWeka) #GOOGLE WEKA AND LEARN ABOUT IT
mushroom_JRip <- JRip(type ~ ., data = mushrooms)
mushroom_JRip
summary(mushroom_JRip)
```

```{r}
detach("package:RWeka", unload = TRUE) #OneR will not print if RWeka is attached
library(C50)
mushroom_c5rules <- C5.0(type~odor+gill_size,data=mushrooms,rules=TRUE)
summary(mushroom_c5rules)


```
